Soling at the 2000 Summer Olympics

### 2020-07-22 inst/f1points.Rmd
cache for years 2016-9, heavy lifting done in f1points_Omaker.R

```{r,label=showanordertableandhead,echo=FALSE,print=FALSE}
library(hyper2,quietly=TRUE)
top <- 11
H <- ordertable2supp(F1_table_2017[seq_len(top),])
```

@@

```{r,label=printapiechart,echo=FALSE,fig.cap="Maximum likelihood estimates \\label{piechartstrength} of the strengths of the top-ranked 11 drivers in the 2017 Formula 1 season"}
mH <- maxp(H)
mH
pie(maxp(H)[c(1,6,11,2,7,3,8,4,9,5,10)])
```

We see the Hamilton has the largest estimated strength (at about
0.39), followed by Vettel and then Bottas at about 0.2 and 0.13
respectively.  However, it should be noted that a likelihood ratio
test [`samep.test()`, supplied with the package] does not reject the
null that Hamilton and Vettel have the same strength ($H_0\colon
\pham= \pvet$) with a likelihood ratio of $e^{1.5}\simeq 4.48$,
corresponding (by Wilks's theorem) to an asymptotic $p$-value of about
$0.08$.

```{r,label=dosameptest, echo=FALSE, print=FALSE, cache=TRUE}
ignore <- samep.test(H,1:2)
```

@@

```{r, label=calcpoints, echo=FALSE, cache=TRUE}
p <- c(25, 18, 15, 12, 10, 8, 6, 4, 2, 1, 0, 0)
points_real <- p
points_inaugural <- c(8,6,4,3,2)
dp <- ordertable2points(as.ordertable(F1_table_2017[seq_len(top),]),p)
dp <- sort(dp,decreasing=TRUE)
dp[] <- seq_along(dp)
m <- maxp(H)
m <- sort(m,decreasing=TRUE)
m[] <- seq_along(m)
```

@@

```{r,label=likelihoodrankingshower,echo=FALSE}
jj <- mH
jj <- sort(jj,decreasing=TRUE)
jj[] <- seq_along(jj)
jj
```

@@

```{r,label=definemultiplemontecarlo, echo=FALSE}
top <- 11
f2017 <- read.table("formula1_2017.txt",header=TRUE)[seq_len(top),1:20]
m <- maxp(ordertable2supp(as.ordertable(f2017)))

## Now work on synthetic dataset generation:
resampling_multiple <- function(m,f2017,pointslist){
    random_table <- rrank(n=ncol(f2017), p=m)
    rownames(random_table) <- colnames(f2017)
    ## "random_table" is a random table; now calculate lstar and pstar for
    ## likelihood ranks and points ranks respectively:

    l_star <- maxp(ordertable2supp(ranktable_to_ordertable(random_table)))
    l_star[] <- seq_along(l_star)
    ## l_star is the likelihood order statistic.

	goodnesses <- seq_along(pointslist)
	for(i in seq_along(pointslist)){
		p_star <- ordertable2points(as.ordertable(ranktable_to_ordertable(random_table)),pointslist[[i]])
		p_star <- sort(p_star,decreasing=TRUE)
		p_star[] <- seq_along(p_star)
		goodness <- sum(cumprod(names(p_star)==names(l_star)))
		goodnesses[i] <- goodness
		}
	return(goodnesses)
}
pointslist <- list(
	real = points_real,
	inau = points_inaugural,
	top1 = rep(1,1),
	top2 = rep(1,2),
	top3 = rep(1,3),
	top4 = rep(1,4),
	top5 = rep(1,5),
	top6 = rep(1,6),
	top7 = rep(1,7),
	top8 = rep(1,8),
	top9 = rep(1,9),
	topt = rep(1,10),  # effectively punishing the last finisher
   	lin1 = 1,   # same as top1
	lin2 = 2:1,
	lin3 = 3:1,
	lin4 = 4:1,
	lin5 = 5:1,
	lin6 = 6:1,
	lin7 = 7:1,
	lin8 = 8:1,
	lin9 = 9:1,
	lint = 10:1,
	line = 11:1,
	zipf = 1/seq_len(top),
	expa = 1/1.01^seq_len(top),
	expb = 1/1.10^seq_len(top),
	expc = 1/1.20^seq_len(top),
	expd = 1/1.50^seq_len(top),
	exp2 = 1/2^seq_len(top),
	exp3 = 1/3^seq_len(top),
	exp4 = 1/4^seq_len(top),
	fis1 = c(1,1,1),
	fis2 = c(1,1,0.9),
	fis3 = c(1,1,1,0.1),
	fis4 = c(1,1,0.9,0.1),
	fis5 = c(1,1,0.5),
	fis6 = c(1,1,1,0.5),
	fis7 = c(1,1,0.5,0.5)
)
```

```{r,label=multiplemontecarloexecute, cache=TRUE,echo=FALSE}
set.seed(9)
f2017 <- read.table("formula1_2017.txt",header=TRUE)[seq_len(top),1:20]
m <- maxp(ordertable2supp(as.ordertable(f2017)))
OO <- replicate(1000,resampling_multiple(m,f2017,pointslist=pointslist))
```

>>>>

```{r,label=showanordertableandhead,echo=FALSE,print=FALSE}
library(hyper2,quietly=TRUE)
top <- 11
H2016 <- ordertable2supp(F1_table_2016[seq_len(top),])
H2017 <- ordertable2supp(F1_table_2017[seq_len(top),])
H2018 <- ordertable2supp(F1_table_2018[seq_len(top),])
H2019 <- ordertable2supp(F1_table_2019[seq_len(top),])
```

@@

```{r,label=maximumlikelihoodcalculations,cache=TRUE}
mH2016 <- maxp(H2016)
mH2017 <- maxp(H2017)
mH2018 <- maxp(H2018)
mH2019 <- maxp(H2019)
```

```{r,label=printapiechart,echo=FALSE,fig.cap="Maximum likelihood estimates \\label{piechartstrength} of the strengths of the top-ranked 11 drivers in Formula 1 motor racing, seasons 2016-2019"}
mH2016
mH2017
mH2018
mH2019
par(mfrow=c(2,2))
par(mar = c(2,2,2,2)/2)
pie(mH2016[c(1,6,11,2,7,3,8,4,9,5,10)],cex=0.7,main="2016")
pie(mH2017[c(1,6,11,2,7,3,8,4,9,5,10)],cex=0.7,main="2017")
pie(mH2018[c(1,6,11,2,7,3,8,4,9,5,10)],cex=0.7,main="2018")
pie(mH2019[c(1,6,11,2,7,3,8,4,9,5,10)],cex=0.7,main="2019")
```

We see in figure \ref{piechartstrength} that the driver with the
largest estimated strength was Rosberg in 2016 (at about 49\%) and in
years 2017-8-9 was Hamilton (at about 29\% 35\% 50\ respectively).
However, it should be noted that a likelihood ratio test
[`samep.test()`, supplied with the package] does not reject, for
example, the null that Hamilton and Vettel have the same strength in
2017 ($H_0\colon \pham= \pvet$) with a likelihood ratio of
$e^{1.5}\simeq 4.48$, corresponding (by Wilks's theorem) to an
asymptotic $p$-value of about $0.08$.

```{r,label=dosameptest, echo=FALSE, print=FALSE, cache=TRUE}
ignore <- samep.test(H2017,1:2)
```

@@

```{r, label=calcpoints, echo=FALSE, cache=TRUE}
p <- c(25, 18, 15, 12, 10, 8, 6, 4, 2, 1, 0, 0)
points_real <- p
points_inaugural <- c(8,6,4,3,2)
dp <- ordertable2points(as.ordertable(F1_table_2017[seq_len(top),]),p)
dp <- sort(dp,decreasing=TRUE)
dp[] <- seq_along(dp)
m <- maxp(H2017)
m <- sort(m,decreasing=TRUE)
m[] <- seq_along(m)
```

@@

```{r,label=likelihoodrankingshower,echo=FALSE}
jj <- mH2017
jj <- sort(jj,decreasing=TRUE)
jj[] <- seq_along(jj)
jj
```

@@

```{r,label=definemultiplemontecarlo, echo=FALSE}
load("formula1_results_2017.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
```
